---
title: "Simulation and Inferential Data Analysis_Part2"
author: "Cynthia Tang"
date: "May 30, 2019"
output:
  word_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The ToothGrowth data

The data was collected from 60 guinea pigs. Each animal received one of three 
dose levels (0.5, 1, 2 mg/day) of vitamin C through either orange juice (OJ) or ascorbic acid (VC).
```{r ToothGrowth data}
library(ggplot2)
library(datasets)
data(ToothGrowth)
head(ToothGrowth)
str(ToothGrowth)
summary(ToothGrowth)
table(ToothGrowth$dose)
ggplot(data = ToothGrowth, aes(x = as.factor(dose), y = len, fill = supp)) +
        geom_boxplot() +
        labs(title = "Plot1. The tooth lengths affected by supplement types and dose")
```

Based on the Plot1, the tooth length increased as the dose of supplements increased.
Also, the animals exposed to orange juice(OJ) had greater tooth length than those exposed to ascorbic(VC).

## Hypothesis tests

**Question 1: Does the orange juice perform better in increasing tooth length in guinea pigs than ascorbic acid?**  
H0: mu(OJ) - mu(VC) <= 0   
Ha: mu(OJ) - mu(VC) > 0  
We used t-test and assumed unequal variance in two supplement types.  
Here are the test results:

```{r type, echo=FALSE}
q1 <- t.test(len ~ supp, data = ToothGrowth, alternative = c("greater"))
q1
```
**Question 2: Does the the higher dosage perform better in increasing tooth length in guinea pigs than the lower dosage ?**  
H0: mu(lower) - mu(higher) >= 0  
Ha: mu(lower) - mu(higher) < 0   
We used t-test and assumed unequal variance in two supplement dosage.  
Here are the test results:

```{r dosage, echo=FALSE}
subdata1 <- subset(ToothGrowth, dose %in% c(0.5, 1))
subdata2 <- subset(ToothGrowth, dose %in% c(0.5, 2))
subdata3 <- subset(ToothGrowth, dose %in% c(1, 2))
q21 <- t.test(len ~ dose, data = subdata1, alternative = c("less"))
q22 <- t.test(len ~ dose, data = subdata2, alternative = c("less"))
q23 <- t.test(len ~ dose, data = subdata3, alternative = c("less"))
data.frame("dosage0.5vs1" = q21$p.value, "dosage0.5vs2" = q22$p.value, "dosage1vs2" = q23$p.value)
```

## Conclusions

The p-value for question 1 was equal to `r q1$p.value` and the p-value for question 2 was
`r q21$p.value`, `r q22$p.value`, `r q23$p.value` for 0.5 mg/day vs 1 mg/day, 0.5 mg/day vs 2 mg/day, and 1 mg/day vs 2mg/day respectively. 

We set type I error (alpha) equal to 0.05, then we rejected the null hypothesis for both
questions and concluded that the orange juice perform better in increasing tooth length in guinea pigs than ascorbic acid and the higher dosage perform better than the lower dosage.

```{r backup, include=FALSE}
# plotting the hist of len
# ggplot(data = ToothGrowth,aes(x = len)) +
#         geom_histogram(binwidth = 2, color = "black") +
#         labs(title = "The distribution of tooth lengths")
# 
# ggplot(data = ToothGrowth,aes(x = len, fill = supp)) +
#         geom_histogram(binwidth = 2, color ="black") +
#         facet_grid(supp~dose) +
#         labs(title = "The distribution of tooth lengths by supplement types and dose")
# 
# ggplot(data = ToothGrowth, aes(x = len)) +
#         geom_histogram(binwidth = 2, color ="black") +
#         facet_grid(.~dose) +
#         labs(title = "The distribution of tooth lengths by dose")
# 
# ggplot(data = ToothGrowth, aes(x = len, fill = supp)) +
#         geom_histogram(binwidth = 2, color ="black") +
#         facet_grid(.~supp) +
#         labs(title = "The distribution of tooth lengths by supplement types")

```

